An efficient approach to approximating the pair distribution function of the inhomogeneous hard-sphere fluid

### 2014-09-02 papers/pair-correlation/paper.tex
polish off end of Section IV Tested on bennet. Took 0:19.21.

The CVA requires the radial distribution function expressed as a
function of $r$ and $g_\sigma$.  \fixme{One approach to constructing
  $g(r;g_\sigma)$ is to compute $n$ from $g_\sigma$ (using e.g. the
  Carnahan-Starling formula for $g_\sigma$) and then find $g(r;n)$
  using either some approximate functional form (e.g. Percus-Yevick)
  or Monte Carlo simulation results.}  \newstuff{In this paper we
  construct a functional form for $g(r, g_\sigma)$ that allows for
  greater computational efficiency. In section \ref{sec:efficient} we
  discuss the general form that allows for this efficiency and in
  section \ref{sec:separable-fit} we detail the fit to Monte Carlo
  results.}  When evaluating the approximations of earlier works
described in Section~\ref{sec:gn} \newstuff{for comparison}, we use interpolated Monte Carlo
results for $g(r;n)$.

@@

\section{Making the CVA efficient}
\label{sec:efficient}
The existing approaches to approximating the pair distribution
function outlined in Section~\ref{sec:gn} have not been widely used in
the construction of density functionals based on thermodynamic
perturbation theory, largely due to their computational complexity.
While our CVA provides only an incremental improvement in accuracy,
its construction enables significant gains in computational
efficiency, allowing for practical application in density functionals.
We achieve this gain by developing a \emph{separable} fit to the radial
distribution function of the hard-sphere fluid (see
Section~\ref{sec:separable-fit} for details).  This separable fit is
of the form
\begin{align}
  g(r; g_\sigma) &= \sum_{i} a_i(r) b_i(g_\sigma)
  \\\text{\newstuff{removed subscript $_{12}$.}}
\end{align}
where the notable aspect is that the radial distribution function is
written as a sum of terms that are each a simple product of a function
of radius with a function of $g_\sigma$.  This enables us to write
integrals---such as Eq.~\ref{eq:mean-field}---that are linear in the
pair distribution function as a summation of fixed-kernel
convolutions, which may be efficiently computed using Fast Fourier
Transforms (FFTs) in $\mathcal{O}(N\log N)$ time.

>>>>

The CVA requires the radial distribution function expressed as a
function of $r$ and $g_\sigma$.
%
%% \fixme{One approach to constructing $g(r;g_\sigma)$ is to compute $n$
%%   from $g_\sigma$ (using e.g. the Carnahan-Starling formula for
%%   $g_\sigma$) and then find $g(r;n)$ using either some approximate
%%   functional form (e.g. Percus-Yevick) or Monte Carlo simulation
%%   results.}
%
We construct a functional form for $g(r, g_\sigma)$ that allows for
improved computational efficiency.  We introduce the general form that
allows for this efficiency in Section~\ref{sec:efficient}, and we
detail our specific approximation for $g(r,g_\sigma)$ that uses this
general form in Section~\ref{sec:separable-fit}.
%
%% When evaluating the approximations of earlier works described in
%% Section~\ref{sec:gn} \newstuff{for comparison}, we use interpolated
%% Monte Carlo results for $g(r;n)$.

@@

\section{Making the CVA efficient}
\label{sec:efficient}
The existing approaches to approximating the pair distribution
function outlined in Section~\ref{sec:gn} have not been widely used in
the construction of density functionals based on thermodynamic
perturbation theory, largely due to their computational complexity.
While our CVA provides only an incremental improvement in accuracy,
its construction enables significant gains in computational
efficiency, allowing for practical application in density functionals.
We achieve this gain by developing a \emph{separable} fit to the radial
distribution function of the hard-sphere fluid (see
Section~\ref{sec:separable-fit} for details).  This separable fit is
of the form
\begin{align}
  g(r; g_\sigma) &= \sum_{i} a_i(r) b_i(g_\sigma)
\end{align}
where the notable aspect is that the radial distribution function is
written as a sum of terms that are each a simple product of a function
of radius with a function of $g_\sigma$.  This enables us to write
integrals---such as Eq.~\ref{eq:mean-field}---that are linear in the
pair distribution function as a summation of fixed-kernel
convolutions, which may be efficiently computed using Fast Fourier
Transforms (FFTs) in $\mathcal{O}(N\log N)$ time.
